---
title: "Given names"
author: "Florian Gaudin-Delrieu"
date: "21 mars 2016"
output: html_document
---

## About 

The idea is to study the data about given names from [wikipedia.org](http://wikipedia.org), and in particular, what are the occupations of people with that given name.
It turns out that scrapping wikipedia, getting the first names and the occupation would be difficult, as the data is unstructured. Hopefully, the [wikidata project](http://wikidata.org) exists. It gives access to structured data, so we will use that to get the informations we want.

```{r Needed libraries, message=FALSE}
library(ggplot2)
library(WikidataR)
library(dplyr)
library(tidyr)
library(magrittr)
library(ggthemes)
library(plotly)
library(SPARQL)
library(wordcloud)
library(stringr)
theme_set(theme_minimal(12))
```

## Retrieving informations

We will get the item reference to the given name we want to study (it starts by a Q and is not meaningful). Then we can use that item to feed a SPARQL query that will give us the dataset.

### Getting the item id

We want to get informations about one first name, let's say "Marcel". We will use the find_item function from the WikidataR library.

```{r Getting the list}
liste <- find_item("Florian") #Limit is set by default to 10 results, it should be enough here.
liste
```

Now we have a list with 10 results, and we want to extract the id of the first entry with "given name", or the first entry. I wrapped these steps into a function.

```{r Getting the id}
selectionnerID<-function(nom){
  liste <- find_item(nom)
  id <-""
  for(i in 1:length(liste)){
    if(is.null(liste[[i]]$description)){
      next
    }
    if(grepl("given name",liste[[i]]$description)){ 
      id <- liste[[i]]$id
        break # we break on the first given name we find
    }
  }
  # If we haven't found given name, then we take the first id
  if(id==""){
    id<-liste[[1]]$id
  } 
  return(id)
}
id<-selectionnerID("Florian")
get_item(id) #Sanity check
```

### Modifying the query

We will query wikidata with a query I came up with, and modifying the given name id we got. The place where to replace is given by **REPLACE_ID**, so that's what we will be looking for to replace.  
The parameters are :

* __P735__ for given name
* __P106__ for occupation
* __P569__ for the date of birt
* __P27__ for the country
* Labels are the french ones, with a fallback in english.

You can find the parameters code using `find_property("parameter")` from WikidataR.

```{r}
endpoint <- "https://query.wikidata.org/bigdata/namespace/wdq/sparql"
#endpoint <- "https://query.wikidata.org/sparql"

prefix<-c("wd","<http://www.wikidata.org/entity/>",
          "wdt", "<http://www.wikidata.org/prop/direct/>",
          "wikibase","<http://wikiba.se/ontology#>")

generic_query <-"SELECT ?item ?itemLabel ?occupationLabel ?paysLabel ?annee
WHERE
{
?item wdt:P735 wd:REPLACE_ID .
?item wdt:P106 ?occupation .
OPTIONAL {?item wdt:P569 ?anneeN} .
OPTIONAL {?item wdt:P27 ?pays} .
BIND(YEAR(?anneeN) as ?annee) .	
SERVICE wikibase:label { bd:serviceParam wikibase:language \"fr,en\" }
}
ORDER BY DESC (?annee)"

query<-sub("REPLACE_ID",id,generic_query)
```

### Executing the request

We will use the SPARQL library to get the dataset.

```{r Query}
source("mySPARQL.R")
res <- mySPARQL(endpoint,query,ns=prefix,format = "xml")
resultats<- res$results
summary(resultats)
glimpse(resultats)
```

### Cleaning the data

We need to format the data we got. The selectionnerNom function will fetch the middle part of the returned column, which is in the form `"GivenName LastName"@fr`. So we will apply this function to the "labels" columns (itemLabel, occupationLabel and paysLabel), and then convert those columns to factors.

```{r Cleaning the data}
selectionnerNom<-function(x,colonne,nom) {
  a<-x %>%
    separate_(colonne,c("debut",nom,"fin"),sep="\"",remove=TRUE) %>%
    select(-debut,-fin)
  a[,nom]<-factor(a[,nom])
  return(a)
}

resultats<-selectionnerNom(resultats,"itemLabel","nom")
resultats<-selectionnerNom(resultats,"occupationLabel","metier")
resultats<-selectionnerNom(resultats,"paysLabel","pays")
resultats<-tbl_df(resultats)
resultats
```


```{r Word Cloud}
library(wordcloud)
tousLesMetiers<-resultats %>%
  group_by(metier) %>%
  count(metier)

set.seed(1)
wordcloud(tousLesMetiers$metier,tousLesMetiers$n,scale=c(3,0.25),min.freq=2,colors=brewer.pal(9,"OrRd")[c(5,6,7,8,9)],random.order = FALSE,rot.per=0.3)
```

Some descriptions are quite long, and sometimes the most occuring profession wouldn't fit on the wordcloud. I made the function that adds a newline character near the middle of the descriptions, if there is a space or a dash.

```{r Cutting words for better wordcloud printing}
couperMot<-function(mot){
  #coupe une chaine de caractère sur l'espace le plus proche du milieu
  z=as.character(mot)
  mi=ceiling(str_length(z)/2)
  a<-str_locate_all(z,"[ -]")[[1]]
  if (length(a)==0){ #S'il n'y a pas d'espace ou tiret
    return(z)
  }
  else { #on prend le premier espace après le milieu
    b<-a[(a-mi>0)[,1],1][1]
    if (length(b)==0){ #Si les espaces ou tiret sont avant le mileu
      b<-a[1,1]
    }
    str_sub(z,b,b)<-"\n"
    return(z)
  }
}
tousLesMetiers$new<-mapply(couperMot,tousLesMetiers$metier)
wordcloud(tousLesMetiers$new,tousLesMetiers$n,scale=c(3,0.25),min.freq=2,colors=brewer.pal(9,"OrRd")[c(5,6,7,8,9)],random.order = FALSE,rot.per=0.3)
```


```{r Wrapping it all together}

gettingData<-function(prenom){
  id<-selectionnerID(prenom)
  
  #Building the query
  endpoint <- "https://query.wikidata.org/bigdata/namespace/wdq/sparql"

  prefix<-c("wd","<http://www.wikidata.org/entity/>",
            "wdt", "<http://www.wikidata.org/prop/direct/>",
            "wikibase","<http://wikiba.se/ontology#>")

  generic_query <-"SELECT ?item ?itemLabel ?occupationLabel ?paysLabel ?annee
  WHERE
  {
  ?item wdt:P735 wd:REPLACE_ID .
  ?item wdt:P106 ?occupation .
  OPTIONAL {?item wdt:P569 ?anneeN} .
  OPTIONAL {?item wdt:P27 ?pays} .
  BIND(YEAR(?anneeN) as ?annee) .	
  SERVICE wikibase:label { bd:serviceParam wikibase:language \"fr,en\" }
  }
  ORDER BY DESC (?annee)"
  
  query<-sub("REPLACE_ID",id,generic_query)
  
  #Getting te data
  results<-mySPARQL(endpoint,query,ns=prefix,format = "xml")$results
  
  #Cleaning the data
  results<-selectionnerNom(results,"itemLabel","nom")
  results<-selectionnerNom(results,"occupationLabel","metier")
  results<-selectionnerNom(results,"paysLabel","pays")
  
  return(results)
}

occupationWordcloud<-function(results){
  everyOccupation<-results %>% 
    group_by(metier) %>% 
    count(metier)
  everyOccupation$occ<-mapply(couperMot,everyOccupation$metier)
  set.seed(14)
  wordcloud(everyOccupation$occ,everyOccupation$n,
            scale=c(3,0.25),min.freq=2,
            colors=brewer.pal(6,"Dark2"),
            random.order = FALSE,rot.per=0.3)
}

occupationGraph<-function(results){
  prenom<-word(results$nom[1])
  ggplotly(ggplot(results %>% distinct(item),aes(x=annee))+
    geom_histogram(binwidth = 10,aes(fill=pays))+
    ggtitle(paste("Répartition des", prenom,"dans Wikidata"))+
    ylab("Nombre")+
    xlab("Année de naissance"))
}

```

```{r Examples}

florian<-gettingData("Florian")
occupationWordcloud(florian)
occupationGraph(florian)

```

